---
title: 'Exponential Distribution in R - A research for Statistical inference project
  (Coursera Data Science Specializacion 2017) '
author: "Antonio Marquez Palacios"
date: "December 22, 2017"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synopsis

Normal distribution may be enough to model solutions for many problems in sciences and engineering, but in some special cases there is a need of a different types of density functions, such as the gamma distribution and the exponential distribution. 

As part of the Statitical Inferente course from the Data Science Specialization in coursera.org, across this document we will research some of the properties of the latter, exponential distribution, which is a special case of the gamma distribution, and compare with the Central Limit Theorem. 

## Simulations
R provides a way to do exponential simulations by calling the rexp(n, rate = 1) function:
- n number of observations. If length(n) > 1, the length is taken to be the number required. For this document n = 40
- rate vector of rates (lambda). For this document labmda = 0.2 will be used



Let's do a single simulation to see how rexp() works:
```{r echo=TRUE}
rexp(40, 0.2)
```
It can be seen that 40 values were simulated with a rate of 0.2.

Now, let's simulate 10000 thousand of exponential distribution samples, for this lets define the following:


Let's consider exponential_simulation as our population, from which will create 10,000 samples (no replacement). This will give us the theoretical mean  (u0), theoretical standard deviation and theoretical variance:

```{r}
library(ggplot2)
n      <- 40
lambda <- 0.2

exponential_simulation <- rexp(n, lambda)
u0    <- mean(exponential_simulation)
sd0   <- sd(exponential_simulation)
var0  <- var(exponential_simulation)
u0
sd0
var0
```
Now proceed to simulate 10,000 samples:

```{r}
B      <- 10000
resamples <- matrix(sample(exponential_simulation, n*B, replace = TRUE), B, n)
resampledMeans <- apply(resamples, 1, mean)
```
Check for the sample mean, sample standard deviation and sample variance using the means:

```{r}
ua    <- mean(resampledMeans)
sda   <- sd(resampledMeans)
vara  <- var(resampledMeans)
ua
sda
vara
```

Let's create a Data Frame to easily compare our theoretical vs sample values

## Comparative between theoretical and Sample values

```{r}
theoVSsampleDF <- data.frame(u = c(u0,ua), sd=c(sd0, sda), var=c(var0, vara))
theoVSsampleDF
```
From the table above it can be seen that u0 is of the same order or ua, they look pretty close to each.
Standard deviations and Variances differences are big


Plot the samples

```{r}
h <- hist(resampledMeans, col="red", xlab="means", main="Means of resampled exponential distribution simulation")
xdnorm <- seq(min(exponential_simulation*10000), max(exponential_simulation*10000), length=1000)
ydnorm <- dnorm(xdnorm, mean=u0*10000, sd=sd0*10000)
ydnorm <- ydnorm * diff(h$mids[1:2])*length(exponential_simulation)
lines(xdnorm, ydnorm, col="black", lwd=2)
```
