---
title: "Basic Inferential Data Analysis for ToothGrowth Data Set: The Effect of Vitamin C on Tooth Growth in Guinea Pigs"
author: "Antonio Marquez Palacios"
date: "December 24, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synopsis

The response is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, (orange juice or ascorbic acid (a form of vitamin C and coded as VC).

## Exploratory analysis

To Load the ToothGrowth data set, first import the datasets library
```{r}
library(datasets)
library(ggplot2)
```

Now let's proceed with some basic exploratory data analysys. Let's do a summary:

```{r}
summary(ToothGrowth)
```
From summary it can be seen the following:
dose range is [0.5,2.0]. Will use these values for confidence intervals. Note here that values with 1.5 are not present in the dose observations
len range is [4.20, 33.90]
we know that len is a function of dose and supp len(dose,supp), so let's see how they relate in a plot:

```{r}
ggplot(ToothGrowth, aes(x=factor(dose), y=len, fill=factor(dose))) + geom_boxplot() + facet_wrap(~supp)+xlab("dose") + ylab("len") + ggtitle("Tooth Growth") + labs(fill="dose")
	
```

## Hypothesis testing

To do the Hypothesis testing, lets assume the following intervals for dose
dose = [0.5, 1.0], dose=[1.0, 2.0], dose=[0.5, 2.0]

```{r}
ToothGrowth_05_10 <- ToothGrowth[ToothGrowth$dose %in% c(0.5,1.0),]
ToothGrowth_10_20 <- ToothGrowth[ToothGrowth$dose %in% c(1.0,2.0),]
ToothGrowth_05_20 <- ToothGrowth[ToothGrowth$dose %in% c(0.5,2.0),]
```

and do the test for each interval:
```{r}
t.test(data=ToothGrowth_05_10, len ~ dose, paired=FALSE)
t.test(data=ToothGrowth_10_20, len ~ dose, paired=FALSE)
t.test(data=ToothGrowth_05_20, len ~ dose, paired=FALSE)
```


## Conclusions

1. Providing a supplement do help on tooth growth, as it can be seen in the box plot above. 
2. Supplement type is not a factor on tooth growth, since we can observe growth in both groups

